Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-12 vesicle_code/docs/adhesion/adhesion.tex
working on revisions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, pass one another,
and then separate.  However, in the presence of adhesion, the vesicles
can form a doublet for certain values of the separation distance
$\delta$, Hamaker constant $\mathcal{H}$, shear rate $\chi$, and reduced
area $\Delta A$.  At all reduced areas, the vesicles have a length of
$2\pi$ to be consistent with simulations in previous sections.

Figures~\ref{fig:doublet090-weakAdhesion}
and~\ref{fig:doublet090-strongAdhesion} show snapshots of two vesicles
that have formed a doublet and the color coding is the tension along the
vesicle.  Each of the vesicles have reduced area $\Delta A = 0.9$, shear
rate $\chi = 0.5$, separation distance $\delta = 0.4$, and Hamaker
constant $\mathcal{H} = 0.7$ (Figure~\ref{fig:doublet090-weakAdhesion})
and $\mathcal{H} = 2.1$ (Figure~\ref{fig:doublet090-strongAdhesion}).
Similar to the quiescent example, the membrane tension is negative in
the contact region indicating that the membrane is being compressed when
the adhesive force is strongest.  For both Hamaker constants, the
vesicles in the doublet move in tandem with a tank-reading motion that
is much slower than the tank-treading motion that occurs in the absense
of ahdesion.

@@

In Figure~\ref{fig:sflow_distance}(a), we plot the minimum distance
between the two vesicles as a function of time for various Hamaker
constants.  For Hamaker constants less than a critical value,
$\mathcal{H}_c$, the adhesive force is not strong enough to form a
doublet, and the distance between the vesicles grows with time (red
curves).  However, when $\mathcal{H} > \mathcal{H}_c$, the adhesive
force is sufficiently strong to bind the vesicles into a doublet, and
then the doublet undergoes a periodic motion (blue curves).  In
Figure~\ref{fig:sflow_distance}(b), we plot the time required for the
doublet to make a complete revolution for three different shear rates.
For a fixed Hamaker constant, as expected, the period decreases for
larger shear rates.  More interestingly, at a fixed shear rate, the
period depends non-monotonically on the Hamaker constant.  When the
Hamaker constant is first sufficiently large to form a doublet, the
individual vesicles undergo a sliding motion as they pass one another,
and the contact region between the vesicles moves along the vesicle
throughout the simulation (Figure~\ref{fig:doublet090-weakAdhesion}).
However, at larger Hamaker constants, once the contact region is formed,
the doublet maintains the same structure and undergoes a tumbling
dynamic similar to a Jeffery orbit.  The simulations illustrated in
Figures~\ref{fig:doublet090-weakAdhesion}
and~\ref{fig:doublet090-strongAdhesion} are chosen so that the doublet
has roughly the same period, but the structure of the doublet is
fundamentally different for these different Hamaker constants.

To further understand the behavior of the Jeffery orbit, we define an
effective aspect ratio $\lambda_{\mathrm{eff}}$ which satisfies
\begin{align*}
  \frac{2\pi}{\chi}\left(\lambda + \lambda^{-1} \right) = p,
\end{align*}
where $p$ is the period of the doublet.  The effective aspect ratio is
the aspect ratio of an ellipse in a shear flow that is undergoing a
Jeffery orbit with period $p$.  For the Hamaker constant $\mathcal{H} =
2$, the shear rates, periods, and effective aspect ratio are summarized
in Table~\ref{tbl:aspectRatio}.  Since the effective aspect ratio is
comparable for all three shear rates, this indicates that the dynamics
of a vesicle doublet in a shear flow resembles a rigid body undergoing a
Jeffery orbit.
\begin{table}[htp]
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 38.06 & 2.92 \\
  0.75 & 24.81 & 2.85 \\
  1.00 & 19.47 & 3.00 
\end{tabular}
\todo[inline]{Update these values with new periods.  Maybe need to
consider more than one Hamaker constant}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
Hamaker constant $\mathcal{H} = 2$, and separation distance $\delta =
0.4$.  The effective aspect ratios are comparable for all three shear
rates.}
\end{table}
\todo[inline]{period results with $\delta$ = 0.1, 0.2, 0.3, 0.4 so that
we can simply say how it affects the period}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The initial placement of the vesicles
is chosen so that the flow dives them towards one another.  In the
absence of adhesion, once the vesicles are sufficiently close, they are
deflected to opposite sides of the $y$ axis, pass one another, and then
separate.  However, in the presence of adhesion, the vesicles can form a
doublet for certain values of the separation distance $\delta$, Hamaker
constant $\mathcal{H}$, shear rate $\chi$, and reduced area $\Delta A$.
At all reduced areas, the vesicles have a length of $2\pi$ to be
consistent with simulations in previous sections.

Figures~\ref{fig:doublet090-weakAdhesion}
and~\ref{fig:doublet090-strongAdhesion} show snapshots of two vesicles
that have formed a doublet and the color coding is the tension along the
vesicle.  Each of the vesicles have reduced area $\Delta A = 0.9$, shear
rate $\chi = 0.5$, separation distance $\delta = 0.4$, and Hamaker
constant $\mathcal{H} = 0.7$ (Figure~\ref{fig:doublet090-weakAdhesion})
and $\mathcal{H} = 2.1$ (Figure~\ref{fig:doublet090-strongAdhesion}).
Similar to the quiescent example, the membrane tension is negative in
the contact region indicating that the membrane is being compressed when
the adhesive force is strongest.  The two values for the Hamaker
constant are chosen since the vesicles dynamics have similar periods of
$t^* \approx 42$.  For both the Hamaker constants, the vesicles in the
doublet move in tandem with a tank-reading motion that is much slower
than the tank-treading motion that occurs in the absense of ahdesion.
However, the dynamics of a single period for theses two Hamaker
constants differ significantly.  For the weaker adhesive force, the
individual vesicles undergo a sliding motion as they pass one another,
and the contact region between the vesicles moves along the vesicle
throughout the simulation (Figure~\ref{fig:doublet090-weakAdhesion}).
However, at larger Hamaker constants, once the contact region is formed,
the doublet maintains the same structure and undergoes a tumbling
dynamic similar to a Jeffery orbit.

@@

To further understand the nature of the doublet,
Figure~\ref{fig:sflow_distance}(a) shows the minimum distance between
the two vesicles as a function of time.  For Hamaker constants less than
a critical value, $\mathcal{H}_c$, the adhesive force is not strong
enough to form a doublet, and the distance between the vesicles grows
with time (red curves).  However, when $\mathcal{H} > \mathcal{H}_c$,
the adhesive force is sufficiently strong to bind the vesicles into a
doublet, and then the doublet undergoes a periodic motion (blue curves).
In Figure~\ref{fig:sflow_distance}(b), we plot the time required for the
doublet to make a complete revolution for three different shear rates.
Initially, as the shear rate increases, the period decreases.  However,
the transition from the weakly adhesive doublet
(Figure~\ref{fig:doublet090-weakAdhesion} to the stronger adhesive
doublet (Figure~\ref{fig:doublet090-strongAdhesion}) results in an
increase in the period.  However, once the Hamaker constant is
sufficiently strong that the contact region remains fixed throughout the
dynamics, the period of the doublet again decreases as the shear rate
increases.

Especially at the large Hamaker constants, the dynamics of the doublet
resemble those of a Jeffery orbit.  An ellipse with aspect ratio
$\lambda$ undergoes a periodic tumbling motion with
period~\cite{jef1922}
\begin{align}
  p = \frac{2\pi}{\chi}\left(\lambda + \lambda^{-1} \right).
  \label{eqn:JefferyPeriod}
\end{align}
Instead of measuring the aspect ratio of a doublet in a flow with shear
rate $\chi$, we use the computed period to define an effective aspect
ratio $\lambda_{\mathrm{eff}}$ that satisfies~\eqref{eqn:JefferyPeriod}.
Therefore, the effective aspect ratio is the aspect ratio of an ellipse
in a shear flow that is undergoing a Jeffery orbit with period $p$.  For
the Hamaker constants $\mathcal{H} = 1.5$ and $\mathcal{H} = 3.5$, the
shear rates, periods, and effective aspect ratio are summarized in
Table~\ref{tbl:aspectRatio}.

Since the effective aspect ratio is
comparable for all three shear rates, this indicates that the dynamics
of a vesicle doublet in a shear flow resembles a rigid body undergoing a
Jeffery orbit.
\begin{table}[htp]
\begin{minipage}{0.45\textwidth}
  $\mathcal{H} = 1.5$ \\
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
%  0.50 & 38.06 & 2.92 \\
%  0.75 & 24.81 & 2.85 \\
%  1.00 & 19.47 & 3.00 
  0.50 & 36.34 & 2.49 \\
  0.75 & 25.86 & 2.72 \\
  1.00 & 20.73 & 2.96 
\end{tabular}
\end{minipage}
%\hfill
\begin{minipage}{0.45\textwidth}
  $\mathcal{H} = 3.5$ \\
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 28.46 & 1.66 \\
  0.75 & 22.01 & 2.17 \\
  1.00 & 17.91 & 2.44 
\end{tabular}
\end{minipage}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
separation distance $\delta = 0.4$, and 
Hamaker constants $\mathcal{H} = 1.5$ and $\mathcal{H} = 3.5$.  The
effective aspect ratios depend on both the Hamaker constant and the
shear rate, indicating that the reported Hamaker constants, the dynamics
of the doublet differ from that of a Jeffery orbit.}
\end{table}
\todo[inline]{period results with $\delta$ = 0.1, 0.2, 0.3, 0.4 so that
we can simply say how it affects the period}
